qcircuit 2.5.3 Tutorial

### 2017-10-16 CO681-QuantumInformationProcessing/assignments/assignment_02.tex
q3
Fixed

$$TXT^{\dagger} = \begin{bmatrix} 1 & 0 \\ 0& \zeta_8 \end{bmatrix}\begin{bmatrix} 0 & 1 \\ 1 & 0 \end{bmatrix}\begin{bmatrix} 1 & 0 \\ 0& \zeta_8^* \end{bmatrix} = \begin{bmatrix} 1 & 0 \\ 0& \zeta_8 \end{bmatrix}\begin{bmatrix} 0 & \zeta^*_8 \\ 1 & 0 \end{bmatrix} = \begin{bmatrix} 0 & \zeta_8^* \\ \zeta_8 & 0 \end{bmatrix} = -e^{\frac{3}{4}i\pi}X.$$
$$TZT^{\dagger} = \begin{bmatrix} 1 & 0 \\ 0& \zeta_8 \end{bmatrix}\begin{bmatrix} 1 & 0 \\ 0 & -1 \end{bmatrix}\begin{bmatrix} 1 & 0 \\ 0& \zeta_8^* \end{bmatrix} = \begin{bmatrix} 1 & 0 \\ 0& \zeta_8 \end{bmatrix}\begin{bmatrix} 1 & 0 \\ 0 & -\zeta^*_8\end{bmatrix} = \begin{bmatrix} 1 & 0 \\ 0 & -1\end{bmatrix} = Z.$$
Thus $T$ is in the Clifford group on $1$ qubit.

@@

\end{document}

>>>>

$$TXT^{\dagger} = \begin{bmatrix} 1 & 0 \\ 0& \zeta_8 \end{bmatrix}\begin{bmatrix} 0 & 1 \\ 1 & 0 \end{bmatrix}\begin{bmatrix} 1 & 0 \\ 0& \zeta_8^* \end{bmatrix} = \begin{bmatrix} 1 & 0 \\ 0& \zeta_8 \end{bmatrix}\begin{bmatrix} 0 & \zeta^*_8 \\ 1 & 0 \end{bmatrix} = \begin{bmatrix} 0 & \zeta_8^* \\ \zeta_8 & 0 \end{bmatrix}.$$
Such a matrix cannot be written as  $\alpha V$ for phase $\alpha$ and $V \in \{I,X,Y,Z\}$. Thus $T$ is not in Clifford group on $1$ qubit.

@@

We use the same notation for gates as in question $2$. $X_4$ is given by the circuit
 \[ \Qcircuit @C=1em @R=1em {
 & \targ &   \qw & \qw\\
 & \ctrl{-1} & \gate{X} & \qw
}\]
$Z_4$ is given by the circuit
\[ \Qcircuit @C=1em @R=1em {
 & \gate{S} &  \gate{S} & \qw\\
 & \gate{S} & \qw & \qw
}\].
$X_8$ is given by the circuit (using one Toffoli gate)
\[ \Qcircuit @C=1em @R=1em {
 & \targ &  \qw & \qw & \qw & \qw\\
 & \ctrl{-1} & \targ& \qw  & \qw &\qw\\
 & \ctrl{-1} & \ctrl{-1} & \gate{X} &\qw & \qw
}\]
Bonus: Figure $4.9$ of Nielsen and Chuang gives a controlled single-qubit implementation of the Toffoli gate. Since my use of the Toffoli gate has different controls I reflect the picture, giving the following implementation of the above Toffoli gate where the first qubit is controlled by the second and third:

\[ \Qcircuit @C=1em @R=1em {
 & \gate{H} & \targ & \gate{T^\dagger} & \targ & \gate{T} & \targ & \gate{T^\dagger}& \targ & \gate{T} & \gate{H} & \qw & \qw & \qw \\
 & \qw & \ctrl{-1}& \qw  & \qw & \qw & \ctrl{-1} & \qw & \qw & \gate{T^\dagger} & \targ & \gate{T^\dagger} & \targ & \gate{S} &\qw\\
 & \qw & \qw & \qw & \ctrl{-2} & \qw & \qw & \qw & \ctrl{-2} & \qw & \ctrl{-1} & \qw & \ctrl{-1} & \gate{T} & \qw
}\]
$Z_8$ is given by the circuit
\[ \Qcircuit @C=1em @R=1em {
 & \gate{T} &  \gate{T} & \gate{T} & \gate{T} & \qw\\
 & \gate{T} & \gate{T}& \qw  & \qw &\qw\\
 & \gate{T} & \qw & \qw &\qw & \qw
}\]
\end{document}
